### 2016-02-24 ch02-qlogms/excitation-design.tex
Lots of work on CH02

  \paragraph{Contents}
  In \secref{sec:excitation:inputDesign}, we first consider (optimal) input design for an example second-order system.
  Next in \secref{}, \gls{LTI} systems are parametrized to facilitate the input design.
  This will lead to signals with a $1/f$ power spectrum as in \secref{sec:excitation:logarithmic}.
  In \secref{sec:excitation:grids} a comparison of different discrete frequency grids is provided, this is afterwards related to the problem of generating a multisine signal with a $1/f$ power spectral density.
  Next, \secref{sec:excitation:optimAlpha} describes the selection of a suitable frequency ratio $\alpha$ for a logarithmically spaced grid such that constant model quality is obtained.
  The usefulness of the proposed signals is illustrated on simulations in \secref{sec:excitation:simulation} and validated on the measurement of a set of band-pass filters in \secref{sec:excitation:measurement}.
  Conclusions are formulated in \secref{sec:excitation:conclusion}.

  \TODO{somewhere: mention optimality criterion is of lesser importance for a ``good'' signal (optimal in X = good in Y in most cases)}

\section{Input design}
\label{sec:excitation:inputDesign}
For the theoretical analysis, we consider \gls{SISO} \gls{OE} systems, i.e.
\begin{align}
y(t) &=G(s,\theta)u(t)  + e(t),
\end{align}
where $u(t)$ is a known input sequence, $y(t)$ the output sequence and $e(t)$ is zero mean Gaussian white noise.
$G$ is a rational, time invariant transfer function in continuous time. 
For notational simplicity, $s$ is used both as the complex Laplace variable and the differentiation operator; it should be clear from the context which is used. 
The transfer function is parameterized by the vector $\theta$.
These parameters are unknown and hence we are interested in estimating their values.

@@

\begin{remark}
D-optimal input designs are quite popular since this criterion minimizes the volume of  the probability density region of the parameters~\citep[Chapter 6]{Goodwin1977}.
Moreover, the D-optimal input design is independent of the parametrization of the system (as long as the different parametrizations can be related by a bijective function).
\end{remark}

@@

 \subsection{A second order system}
Consider the second order system
\begin{align}
y(t) &= G(s,\theta) u(t) = \frac{1}{\frac{s^2}{\wn^2} + 2\frac{\damping}{\wn}s + 1}u(t),
\label{eq:excitation:2ndOrderSys}
\end{align}
with parameters $\theta=[\wn\;\damping]^{\T}$. The resonance frequency of the system is $\wR \isdef \sqrt{1-\damping^2}\wn$
and the $3\unit{dB}$ bandwidth of the system is $\wdB \isdef 2\damping\wn$.
Since information is local in the frequency domain, the input signal power inside $\wdB$, contributes significantly to the information matrix while input power outside $\wdB$ has a smaller contribution. This can be seen by considering the optimal design for a second order system. From \citep[Example 6.4.5]{Goodwin1977} we have that the D-optimal design for \eqref{eq:excitation:2ndOrderSys} is a single sinusoid at frequency
\begin{align}
\omega^\star =\frac%
{\wn
  \sqrt{
    \left(1-2\damping^2\right)
    +
    \sqrt{\left(1-2\damping^2\right)^2+15}
    }
  }
{\sqrt{5}}.
\label{eq:excitation:optFreq}
\end{align}
For lightly-damped systems, i.e $\damping \ll 1$ , $\omega^\star\approx \wn$, which is inside the bandwidth of the system. 
From \eqref{eq:excitation:optFreq} it is clear that the optimal design depends on the resonance of the system to be identified. 
Input design can be based on a guess of $\wn$, however if this guess is wrong, the resulting estimates may be unacceptable~\citep{Rojas2007}.

@@

Second, we consider the power spectrum of the quasi-logarithmic multisine signal [\cite{Pintelon2012}] given by
\begin{align}
  \phi_u(\omega) &= \frac{1}{M}\sum_{k=1}^{M}{\dirac{\omega - \omega_k}},
\end{align}
where $\omega_{k}/\omega_{k+1} \approx \alpha$ and $\dirac{\omega}$ denotes the Dirac distribution.
Hence, the number of spectral lines inside \wdB is 
\begin{align}
  \floor{
    \log_{\alpha} \frac{1+\damping}{1-\damping} 
  }
  \text{,}
\end{align}
where $\floor{\placeholder}$ is the nearest integer rounded towards $0$. The number of spectral lines is independent of $\wn$.

More generally, a multisine can be defined as:
\begin{equation}
 u \left[ n\right] = \frac{1}{\sqrt{F}}
   \sum_{k=1}^{F} 
     A_k 
     \sin 
       \left(\frac{2\pi n k f_0}{f_s} + \psi_k \right)
  \label{eq:excitation:MS}
\end{equation}
where $A_k$ and $\psi_k$ are its amplitude and phase spectrum; $f_s$ is the
sampling frequency and $f_0 = \frac{f_s}{N}$ the base frequency with $N$ the 
number of points in a period.

To construct a quasi-logarithmic multisine, only frequency lines $f_j$ within the set of available sine components $\left\{f_0, 2 f_0, \ldots, F f_0  \right\}$ are excited for which the ratio of two consecutive frequencies $f_j$ and $f_{j-1}$ is approximately constant: $\frac{f_{j-1}}{f_j} \approx \alpha$.

@@

\subsection{Generic system}
%TODO: explain generic LTI system
Consider the second-order subsystem with multiplicity $\multiplicity$:
\begin{equation}
  G_2(s, \theta) = \frac{K \wn^{2 \multiplicity} \left( s + \wz \right)}{\wz \left(s^2 + 2 \damping \wn s + \wn^2 \right)^{\multiplicity}}
\end{equation}
parametrized in $\theta =  \left[K, \wn, \wz, \damping \right]^{\T}$.
To determine the Fisher Information matrix, the derivative $\PartialDerivative{G(s,\theta)}{\theta}$ is required; its components are given by:
\begin{align}
  \PartialDerivative{G_2}{K} 
    &= 
  \frac{\wn^{2\multiplicity}}{\wz}
  \frac{s + \wz}{\left(s^2 + 2 \damping \wn s + \wn^2\right)^{\multiplicity}}
  \\
  \PartialDerivative{G_2}{\wz} 
    &= 
    \frac{-K \wn^{2\multiplicity}}
              {\wz} 
    \frac{s}
              {\left(s^2 + 2 \damping \wn s + \wn^2\right)^{\multiplicity}} 
  \\
  \PartialDerivative{G_2}{\wn} 
  &= 
  \frac{2 K \wn^{2\multiplicity-1} \multiplicity}
            {\wz} 
  \frac{\left( s + \wz \right) \left( s+ \damping \wn \right)}
            {\left( s^2 + 2 \damping \wn s + \wn^2 \right)^{\multiplicity + 1}}
  \\
  \PartialDerivative{G_2}{\damping}
  &= 
  \frac{-2 K \wn^{2\multiplicity + 1} \multiplicity}
            {\wz}
  \frac{s \left( s + \wz \right)}
             {\left( s^2 + 2 \damping \wn s + \wn^2 \right)^{\multiplicity + 1}}
  \text{.}
\end{align}

The poles of these derivatives coincide with the system poles (possibly with increased multiplicity).
Consequently, this means that for lightly-damped systems ($\damping \ll \sqrt{2}$), these derivatives will exhibit a sharp resonance peak just as the actual sub-system $G_2$ as illustrated in \figref{fig:excitation:derivatives}.
The roll-off near the resonance is determined mainly by the damping $\damping$ and the multiplicity $\multiplicity$ of the sub-system.
As such, the Fisher information matrix is pre-dominantly determined by the input power spectrum near $\wn$.
Note that $\PartialDerivative{G_2}{\wn}$ and $\PartialDerivative{G_2}{\damping}$ have the steepest roll-off near the resonance.
As such, those parameters are the most important for the experiment design.

@@

\subsubsection{Example: Sum of two second order systems}
Consider the system
\begin{align}
G(s,\theta) &= G_1(s,\theta_1) + G_2(s,\theta_2)%\\
%&=\frac{1}{\frac{s^2}{\wn[1]^2} + 2\frac{\damping[1]}{\wn[1]}s + 1} + \frac{1}{\frac{s^2}{\wn[2]^2} + 2\frac{\\damping[2]}{\wn[2]}s + 1},
\end{align}
with $\theta_1=[\wn[1]\;\damping[1]]^{\T}$ and $\theta_2=[\wn[2]\;\damping[2]]^{\T}$.
The information matrix for this system is
\begin{align}
\IF &=
       \begin{bmatrix}
         \partial G_{1,1}  & \partial G_{1,2}\\
         \partial G_{2,1}  & \partial G_{2,2}
       \end{bmatrix}.
\label{eq:excitation:infoEx}
\end{align}
To investigate the decoupling, we study the $(1,1)$ elements of blocks $\partial G_{1,1}$ and $\partial G_{1,2}$ of the information matrix \eqref{eq:excitation:infoEx}, when the input spectrum \eqref{eq:excitation:pink} is used; similar arguments can be made for the other elements of the two matrix blocks.

@@

\begin{figure}
\centering
\setlength\figurewidth{0.68\columnwidth}
\setlength\figureheight{0.68\figurewidth}
\input{\thisDir/figs/coupling.tikz}
\caption{The element with the largest absolute value in the off-diagonal blocks of the normalized information matrix for different dampings.
         The different plots are for different separations of the resonance frequencies, from 4\wdB to 64\wdB.
         The arrow indicates the direction of increasing separation.}
\label{fig:excitation:coupling}
\end{figure}

\section{Methodology} 
\label{sec:excitation:method}

  \subsection{The system under consideration}
  We consider the transfer function models of generic \gls{LTI} systems of arbitrary but finite orders $\order{B} / \order{A}$:
  \begin{align}
    G\left( \Omega, \theta \right) 
    = \frac{B\left( \Omega,\theta \right)}
                  {A\left( \Omega,\theta \right)}
    = \frac{\sum\limits_{i=0}^{\order{B}} b_i \Omega^i}
                  {\sum\limits_{i=0}^{\order{A}} a_i \Omega^i}
    \text{.} 
    \label{eq:excitation:model}
  \end{align}

  The model is parametrized by the vector $\theta = \left[b_0 \,\cdots\, b_{n_b-1} \; a_0 \,\cdots\, a_{n_a-1} \right]^{\T}$ and  is evaluated at different frequencies $\Omega$ in the complex plane.
  For a continuous time model $\Omega = s$, while for the discrete time case $\Omega = z$.

  Such a system can be decomposed into a series of first and second order systems by the partial fraction expansion:
Remember that the rational transfer function of any strictly proper \gls{LTI} system $G$ can be decomposed in its partial fraction form\citep[Appendix]{Oppenheim1996}:
\begin{align}
  G(s) &= 
    \frac{\sum\limits_{i=0}^{\numel{B}} b_i s^i}
              {\sum\limits_{i=0}^{\numel{A}} a_i s^i} 
        =    
   \frac{\sum\limits_{i=0}^{\numel{B}} b_i s^i}
              {\prod\limits_{i=1}^{I_1} \left( s - \pole[i] \right)^{M_{1,i}} 
               \prod\limits_{i=1}^{I_2} \left(s^2 + 2 \damping[i] \wn[i] s + \wn[i]^2 \right)^{M_{2,i}}}
        \\
        &  = 
        \sum_{i=1}^{I_1} 
          \sum_{\multiplicity=1}^{M_{1,i}} 
            \frac{\dc[i]}
                      { \left( s - \pole[i] \right)^{\multiplicity}}
            + 
        \sum_{i=1}^{I_2} 
          \sum_{\multiplicity=1}^{M_{2,i}}
           \frac{\dc[i,\multiplicity] \wn[i]^{2\multiplicity} \left( s + \wz[i] \right)}
                     {\wz[i] \left(s^2 + 2 \damping[i] \wn[i] s + \wn[i]^2 \right)^{\multiplicity}}
        \label{eq:excitation:PFE}
\end{align}
where $I_1$ denotes the number of real-valued poles, $I_2$ denotes the number of complex conjugated pole pairs.
$M_{1,i}$ and $M_{2,i}$ denotes the multiplicity of the corresponding poles, such that $\numel{A} = \sum_i^{I_1} M_{1,i} + 2 \sum_{i}^{I_2} M_{2,i}$.
Each term in \eqref{eq:excitation:PFE} represents a \term{sub-system} of the system $G(s)$.
  Within these sub-systems, we can distinguish first-order subsystems and second-order subsystems, the latter of which describe resonant phenomena.
  Since such resonances are often a limiting factor, we first focus on these resonant sub-systems.
\begin{remark}
  For non-proper systems, one can write the transfer function $G_{\mathrm{non-proper}}(s) = P(s) + G_{\mathrm{proper}}$ where $P(s)$ is a polynomial and $G_{\mathrm{proper}}(s)$ a proper rational function.
  As such, the polynomial $P(s)$ should be added to equation \eqref{eq:excitation:PFE} for non-proper systems.
\end{remark}

  To study the behavior of one sub-system, we inspect a generic second order system $G\left( s \right)$.
  In continuous time such a system can be described by a transfer function of the form
  \begin{equation}
    G\left(s \right) = \frac{\dc}
                            {             \frac{s^2}
                                               {\wn^2} 
                             + 2 \damping \frac{s}
                                               {\wn}
                             + 1
                            }
  \text{.}
  \label{eq:excitation:secondOrderSystemCT}
  \end{equation}

  The DC gain $\dc$, the (relative) damping $\damping$ and the natural pulsation $\wn$ are a more intuitive parametrization for this system than the generic polynomial coefficients $a_i$ and $b_i$ used in \eqref{eq:excitation:model}.
  The corresponding resonance pulsation is $\wR \isdef \wn \sqrt{1-\damping^2}$.
  The $3\unit{dB}$ bandwidth around the resonance is given by $\wdB = 2\damping\wn$ for systems that have a sufficiently low damping.

  When $\damping < \frac{1}{\sqrt{2}}$, such a second order system exhibits a peak in
  the amplitude of the transfer function \citep{Oppenheim1983}:
  \begin{align}
   \left| G\left(\wM \right) \right| & \isdef
          \max_{\omega} \left| G\left( \omega \right)  \right|
        = \frac{\dc}
               {2 \damping \sqrt{1-\damping^2}}
               \label{eq:excitation:peakAmplitude}\\
   \wM &
        = \sqrt{1-2\damping^2} \wn
        \label{eq:excitation:peakFrequency}
    \text{.}
  \end{align}

\subsection{System and noise model}
  In this chapter, an \gls{OE} system setup is used.
  Denote
    $U  \left(\omega\right)$ as the excitation signal,
    $\true{Y}\left(\omega\right)$ the exact, unperturbed system output,
    $E  \left(\omega\right)$ for the output noise source,
    $Y  \left(\omega\right) = \true{Y}\left( \omega \right) + E\left( \omega \right)$ depicts the measured output signal as shown in \figref{fig:excitation:OE}.

  \begin{figure}
    \centering
    \input{\thisDir/figs/oesetup.tikz}
    \caption[Output-error]{Output-error set-up.}
    \label{fig:excitation:OE}
  \end{figure}

  The output noise is assumed to be zero-mean and white with a
  standard deviation of $\sigma_e$ in the time domain or $\sigma_E$ in the frequency domain.

\subsection{Multisine Excitations}
  To excite the system under test we consider a generalization of multisine excitations \citep{Pintelon2012}:
  \begin{definition}[Generalized multisine] \label{def:excitation:generalized-MS}
  A generalized multisine $u(t)$ is a signal consisting of $F$ sine waves with a different frequencies $f_k$, amplitudes $A_k$ and phase shifts $\phi_k$:
  \begin{equation}
    u \left( t\right) = \frac{1}{\sqrt{F}}
   \sum_{k=1}^{F} 
     A_k 
     \sin 
       \left(2\pi f_k t + \phi_k \right)
  \text{.}
  \label{eq:excitation:MultisineCT}
  \end{equation}
  Without loss of generality, one can stipulate that the frequency grid is sorted, i.e. $f_1 < f_2 < \ldots < f_F$.
  \end{definition}

  For \gls{LTI} systems, it is well-known \TODO{refer to eqn. from Larsson} that the information matrix only depends upon the amplitude spectrum of the input and not upon its phase spectrum.
  As such, in this chapter assume random-phase multisines since this is a reasonable choice for systems that are dominantly linear~\TODO{referentie}.
  Hence, $\phi_k$ is the outcome of a uniform random process over $\left[0,2\pi\right[$.
  This leaves two sets of free parameters to create a signal that suits our needs:
    the amplitude of the spectral lines $A_k$ and
    the frequency grid $\left\{f_1, \ldots, f_F \right\}$.
  Note that the $k^{\text{th}}$ term in \eqref{eq:excitation:MultisineCT} is periodic with period $T_k \isdef f_k^{-1}$.

@@

\TODO{cov-matrix uitschrijven}

\TODO{zin loopt niet goed}
  As in \citep{Pintelon2012, Gallager2008, matrixcookbook}, the covariance matrix $\C_{G\left(\omega_{\interest}\right)}$ of the parametric transfer function $G(\theta,\omega)$ can be shown to be approximately equal to
  \begin{align}
    \C_{G\left( \omega_{\interest}\right) } \approx
      \InnerQuadraticForm[\HT]
                         {\jac{\interest}}
                         {\FisherInformation_{\theta}^{-1}}
     \label{eq:excitation:varPeak}
  \end{align}
  with 
  \begin{equation}
    \FisherInformation_{\theta} =
    2
    \real{
          \InnerQuadraticForm[\HT]
                              {\jac{\excited}}
                              {\C_{X}^{-1}}
          }
    \label{eq:excitation:FisherInfo}
  \end{equation}
  the Fisher information matrix \citep{Pintelon2012}.

@@

  \paragraph{Simplification}
  Instead of studying the response of all possible second-order sub-systems for every possible logarithmically spaced multisine-- a five-dimensional problem with parameters $(\omega_0, \alpha, \wn, \damping, \dc)$-- we argue in favor for a simplification of the problem by considering frequency grids with an infinite span.

>>>>

  \paragraph{Contents}
  In \secref{sec:excitation:inputDesign}, we first consider (optimal) input design for an example second-order system.
  Next in \secref{}, \gls{LTI} systems are parametrized to facilitate the input design.
  This will lead to signals with a $1/f$ power spectrum as in \secref{sec:excitation:logarithmic}.
  In \secref{sec:excitation:multisine} a comparison of different discrete frequency grids is provided, this is afterwards related to the problem of generating a multisine signal with a $1/f$ power spectral density.
  Next, \secref{sec:excitation:optimAlpha} describes the selection of a suitable frequency ratio $\alpha$ for a logarithmically spaced grid such that constant model quality is obtained.
  The usefulness of the proposed signals is illustrated on simulations in \secref{sec:excitation:simulation} and validated on the measurement of a set of band-pass filters in \secref{sec:excitation:measurement}.
  Conclusions are formulated in \secref{sec:excitation:conclusion}.

\section{Input design}
\label{sec:excitation:inputDesign}
For the theoretical analysis, we consider \gls{SISO} \gls{OE} systems, i.e.
\begin{align}
y(t) &=G(s,\theta)u(t)  + e(t),
\end{align}
where $u(t)$ is a known input sequence, $y(t)$ the output sequence and $e(t)$ is zero mean Gaussian white noise with variance $\sigma_{e}^2$.
$G$ is a rational, time invariant transfer function in continuous time.
For notational simplicity, $s$ is used both as the complex Laplace variable and the differentiation operator; it should be clear from the context which is used. 
The transfer function is parameterized by the vector $\theta$.
These parameters are unknown and hence we are interested in estimating their values.

@@

\begin{remark}
D-optimal input designs are quite popular since this criterion minimizes the volume of uncertainty ellipsoid of the parameters.
Moreover, the D-optimal input design is independent of the parametrization of the system (as long as the different parametrizations can be related by a bijective function)~\citep[Chapter 6]{Goodwin1977}.
\end{remark}

@@

 \subsection{A second order system}
Consider the second order system
\begin{align}
y(t) &= G(s,\theta) u(t) = \frac{1}{\frac{s^2}{\wn^2} + 2\frac{\damping}{\wn}s + 1}u(t),
\label{eq:excitation:2ndOrderSys}
\end{align}
with parameters $\theta=[\wn\;\damping]^{\T}$. 
The resonance frequency of the system is $\wR \isdef \sqrt{1-\damping^2}\wn$ and the $3\unit{dB}$ bandwidth of the system is $\wdB \isdef 2\damping\wn$.
The $3\unit{dB}$ bandwidth around the resonance is given by $\wdB = 2\damping\wn$ for systems that have a sufficiently low damping.

When $\damping < \frac{1}{\sqrt{2}}$, such a second order system exhibits a peak in the amplitude of the transfer function~\citep{Oppenheim1983}:
  \begin{align}
   \left| G\left(\wM \right) \right| & \isdef
          \max_{\omega} \left| G\left( \omega \right)  \right|
        = \frac{\dc}
               {2 \damping \sqrt{1-\damping^2}}
               \label{eq:excitation:peakAmplitude}\\
   \wM &
        = \sqrt{1-2\damping^2} \wn
        \label{eq:excitation:peakFrequency}
    \text{.}
  \end{align}

Since information is local in the frequency domain, the input signal power inside $\wdB$, contributes significantly to the information matrix while input power outside $\wdB$ has a smaller contribution. This can be seen by considering the optimal design for a second order system. From \citep[Example 6.4.5]{Goodwin1977} we have that the D-optimal design for \eqref{eq:excitation:2ndOrderSys} is a single sinusoid at frequency
\begin{align}
\omega^\star =\frac%
{\wn
  \sqrt{
    \left(1-2\damping^2\right)
    +
    \sqrt{\left(1-2\damping^2\right)^2+15}
    }
  }
{\sqrt{5}}.
\label{eq:excitation:optFreq}
\end{align}
For lightly-damped systems, i.e $\damping \ll 1$ , $\omega^\star\approx \wn$, which is inside the bandwidth of the system. 
From \eqref{eq:excitation:optFreq} it is clear that the optimal design depends on the resonance of the system to be identified. 
Input design can be based on a guess of $\wn$, however if this guess is wrong, the resulting estimates may be unacceptable~\citep{Rojas2007}.

@@

Alternatively, one could also construct a signal with power spectrum concentrated at discrete frequency lines.
Such signals are elaborated in \secref{sec:excitation:method}.

% Second, we consider the power spectrum of the quasi-logarithmic multisine signal [\cite{Pintelon2012}] given by
% \begin{align}
%   \phi_u(\omega) &= \frac{1}{M}\sum_{k=1}^{M}{\dirac{\omega - \omega_k}},
% \end{align}
% where $\omega_{k}/\omega_{k+1} \approx \alpha$ and $\dirac{\omega}$ denotes the Dirac distribution.
% Hence, the number of spectral lines inside \wdB is 
% \begin{align}
%   \floor{
%     \log_{\alpha} \frac{1+\damping}{1-\damping} 
%   }
%   \text{,}
% \end{align}
% where $\floor{\placeholder}$ is the nearest integer rounded downwards. 
% The number of spectral lines is independent of $\wn$.

% More generally, a multisine can be defined as:
% \begin{equation}
%  u \left[ n\right] = \frac{1}{\sqrt{F}}
%    \sum_{k=1}^{F} 
%      A_k 
%      \sin 
%        \left(\frac{2\pi n k f_0}{f_s} + \psi_k \right)
%   \label{eq:excitation:MS}
% \end{equation}
% where $A_k$ and $\psi_k$ are its amplitude and phase spectrum; $f_s$ is the
% sampling frequency and $f_0 = \frac{f_s}{N}$ the base frequency with $N$ the 
% number of points in a period.

% To construct a quasi-logarithmic multisine, only frequency lines $f_j$ within the set of available sine components $\left\{f_0, 2 f_0, \ldots, F f_0  \right\}$ are excited for which the ratio of two consecutive frequencies $f_j$ and $f_{j-1}$ is approximately constant: $\frac{f_{j-1}}{f_j} \approx \alpha$.

@@

@@

\subsubsection{Example: Sum of two second-order systems}
\label{sec:excitation:ex:sumOfSecondOrderSipmle}
Consider the system
\begin{align}
G(s,\theta) &= G_1(s,\theta_1) + G_2(s,\theta_2)%\\
%&=\frac{1}{\frac{s^2}{\wn[1]^2} + 2\frac{\damping[1]}{\wn[1]}s + 1} + \frac{1}{\frac{s^2}{\wn[2]^2} + 2\frac{\\damping[2]}{\wn[2]}s + 1},
\end{align}
with $\theta_1=[\wn[1]\;\damping[1]]^{\T}$ and $\theta_2=[\wn[2]\;\damping[2]]^{\T}$.
The information matrix for this system is
\begin{align}
\IF &=
       \begin{bmatrix}
         \partial G_{1,1}  & \partial G_{1,2}\\
         \partial G_{2,1}  & \partial G_{2,2}
       \end{bmatrix}.
\label{eq:excitation:infoEx}
\end{align}
To investigate the decoupling, we study the $(1,1)$ elements of blocks $\partial G_{1,1}$ and $\partial G_{1,2}$ of the information matrix \eqref{eq:excitation:infoEx}, when the input spectrum \eqref{eq:excitation:pink} is used; similar arguments can be made for the other elements of the two matrix blocks.

@@

\begin{figure}
\centering
\setlength\figurewidth{0.68\columnwidth}
\setlength\figureheight{0.68\figurewidth}
\input{\thisDir/figs/coupling.tikz}
\caption[Relative amplitudes of the off-diagonal blocks of the Fisher information matrix.]{The element with the largest absolute value in the off-diagonal blocks of the normalized information matrix for different dampings.
         The different plots are for different separations of the resonance frequencies, from 4\wdB to 64\wdB.
         The arrow indicates the direction of increasing separation.}
\label{fig:excitation:coupling}
\end{figure}

\section{System parametrization} 
\label{sec:excitation:parametrization}

  We consider the transfer function models of proper \gls{LTI} systems of arbitrary but finite orders $\order{B} / \order{A}$:
  \begin{align}
    G\left( \Omega, \theta \right) 
    = \frac{B\left( \Omega,\theta \right)}
                  {A\left( \Omega,\theta \right)}
    = \frac{\sum\limits_{i=0}^{\order{B}} b_i \Omega^i}
                  {\sum\limits_{i=0}^{\order{A}} a_i \Omega^i}
    \text{.} 
    \label{eq:excitation:model}
  \end{align}

The model is parametrized by the vector $\theta = \left[b_0 \,\cdots\, b_{n_b-1} \; a_0 \,\cdots\, a_{n_a-1} \right]^{\T}$ and  is evaluated at different frequencies $\Omega$ in the complex plane.
For a continuous time model $\Omega = s$, while for the discrete time case $\Omega = z$.
For the sake of notational simplicity, $\Omega = s$ will be used further on, but the results are similar for $\Omega=z$.
Remember that the rational transfer function of any strictly proper \gls{LTI} system $G$ can be decomposed in its partial fraction form~\citep[Appendix]{Oppenheim1996}:
\begin{align}
  G(s) &= 
    \frac{\sum\limits_{i=0}^{\numel{B}} b_i s^i}
              {\sum\limits_{i=0}^{\numel{A}} a_i s^i} 
        =    
   \frac{\sum\limits_{i=0}^{\numel{B}} b_i s^i}
              {\prod\limits_{i=1}^{I_1} \left( s - \pole[i] \right)^{M_{1,i}} 
               \prod\limits_{i=1}^{I_2} \left(s^2 + 2 \damping[i] \wn[i] s + \wn[i]^2 \right)^{M_{2,i}}}
        \\
        &  = 
        \sum_{i=1}^{I_1} 
          \sum_{\multiplicity=1}^{M_{1,i}} 
            \frac{\dc[i]}
                      { \left( s - \pole[i] \right)^{\multiplicity}}
            + 
        \sum_{i=1}^{I_2} 
          \sum_{\multiplicity=1}^{M_{2,i}}
           \frac{\dc[i,\multiplicity] \wn[i]^{2\multiplicity} \left( s + \wz[i] \right)}
                     {\wz[i] \left(s^2 + 2 \damping[i] \wn[i] s + \wn[i]^2 \right)^{\multiplicity}}
        \label{eq:excitation:PFE}
\end{align}
where $I_1$ denotes the number of real-valued poles, $I_2$ denotes the number of complex conjugated pole pairs.
$M_{1,i}$ and $M_{2,i}$ denotes the multiplicity of the corresponding poles, such that $\numel{A} = \sum_i^{I_1} M_{1,i} + 2 \sum_{i}^{I_2} M_{2,i}$.
Each term in \eqref{eq:excitation:PFE} represents a \term{sub-system} of the system $G(s)$.
  Within these sub-systems, we can distinguish first-order subsystems and second-order subsystems, the latter of which describe resonant phenomena.
  Since such resonances are often a limiting factor, we first focus on these resonant sub-systems.
\begin{remark}
  For non-proper systems, one can write the transfer function $G_{\mathrm{non-proper}}(s) = P(s) + G_{\mathrm{proper}}$ where $P(s)$ is a polynomial and $G_{\mathrm{proper}}(s)$ a proper rational function.
  As such, the polynomial $P(s)$ should be added to equation \eqref{eq:excitation:PFE} for non-proper systems.
\end{remark}

Consider the second-order subsystem with multiplicity $\multiplicity$:
\begin{equation}
  G_2(s, \theta) = \frac{K \wn^{2 \multiplicity} \left( s + \wz \right)}{\wz \left(s^2 + 2 \damping \wn s + \wn^2 \right)^{\multiplicity}}
\end{equation}
parametrized in $\theta =  \left[K, \wn, \wz, \damping \right]^{\T}$.
To determine the Fisher Information matrix, the derivative $\PartialDerivative{G(s,\theta)}{\theta}$ is required; its components are given by:
\begin{align}
  \PartialDerivative{G_2}{K} 
    &= 
  \frac{\wn^{2\multiplicity}}{\wz}
  \frac{s + \wz}{\left(s^2 + 2 \damping \wn s + \wn^2\right)^{\multiplicity}}
  \\
  \PartialDerivative{G_2}{\wz} 
    &= 
    \frac{-K \wn^{2\multiplicity}}
              {\wz} 
    \frac{s}
              {\left(s^2 + 2 \damping \wn s + \wn^2\right)^{\multiplicity}} 
  \\
  \PartialDerivative{G_2}{\wn} 
  &= 
  \frac{2 K \wn^{2\multiplicity-1} \multiplicity}
            {\wz} 
  \frac{\left( s + \wz \right) \left( s+ \damping \wn \right)}
            {\left( s^2 + 2 \damping \wn s + \wn^2 \right)^{\multiplicity + 1}}
  \\
  \PartialDerivative{G_2}{\damping}
  &= 
  \frac{-2 K \wn^{2\multiplicity + 1} \multiplicity}
            {\wz}
  \frac{s \left( s + \wz \right)}
             {\left( s^2 + 2 \damping \wn s + \wn^2 \right)^{\multiplicity + 1}}
  \text{.}
\end{align}

The poles of these derivatives coincide with the system poles (possibly with increased multiplicity).
Consequently, this means that for lightly-damped systems ($\damping \ll \sqrt{2}$), these derivatives exhibit a sharp resonance peak at the same frequency as the actual sub-system $G_2$, see also \figref{fig:excitation:derivatives}.
The roll-off near the resonance is determined mainly by the damping $\damping$ and the multiplicity $\multiplicity$ of the sub-system.
As such, the Fisher information matrix is pre-dominantly determined by the input power spectrum near $\wn$.
Note that $\PartialDerivative{G_2}{\wn}$ and $\PartialDerivative{G_2}{\damping}$ have the steepest roll-off near the resonance.
As such, those parameters are the most important for the experiment design.

Moreover, the results from \secref{sec:excitation:ex:sumOfSecondOrderSipmle} for the sum of simplified second order systems, i.e. equation \eqref{eq:excitation:sumSys}, extend directly to the generic \gls{LTI} system in \eqref{eq:excitation:PFE} for the second-order sub-systems.

\paragraph{System and noise model}
  In the remainder of this chapter, an \gls{OE} setting is used.
  Denote
    $U  \left(\omega\right)$ as the excitation signal,
    $\true{Y}\left(\omega\right)$ the exact, unperturbed system output,
    $E  \left(\omega\right)$ for the output noise source,
    $Y  \left(\omega\right) = \true{Y}\left( \omega \right) + E\left( \omega \right)$ depicts the measured output signal as shown in \figref{fig:excitation:OE}.
    The output noise is assumed to be zero-mean and white with a standard deviation of $\sigma_e$ in the time domain or $\sigma_E$ in the frequency domain.

  \begin{figure}[h]
    \centering
    \input{\thisDir/figs/oesetup.tikz}
    \caption[Output-error]{Output-error set-up.}
    \label{fig:excitation:OE}
  \end{figure}

\section{Multisine Excitations}
\label{sec:excitation:multisine}
In contrast to signal with a continuous $1/f$ \gls{PSD} presented in \secref{sec:excitation:logarithmic}, similar results can be obtained by using signals that have a discrete frequency grid.
Such signals are presented in this section.

  To excite the system under test we consider a generalization of multisine excitations \citep{Pintelon2012}:
  \begin{definition}[Generalized multisine] \label{def:excitation:generalized-MS}
  A generalized multisine $u(t)$ is a signal consisting of $F$ sine waves with a different frequencies $f_k$, amplitudes $A_k$ and phase shifts $\phi_k$:
  \begin{equation}
    u \left( t\right) = \frac{1}{\sqrt{F}}
   \sum_{k=1}^{F} 
     A_k 
     \sin 
       \left(2\pi f_k t + \phi_k \right)
  \text{.}
  \label{eq:excitation:MultisineCT}
  \end{equation}
  Without loss of generality, one can stipulate that the frequency grid is sorted, i.e. $f_1 < f_2 < \ldots < f_F$.
  \end{definition}

  For \gls{LTI} systems, it is well-known that the information matrix only depends upon the amplitude spectrum of the input and not upon its phase spectrum as indicated in \eqref{eq:excitation:IF}.
  As such, the results presented in this chapter also do not depend on the particular choice of the phase spectrum.
  This means a user is free to make an appropriate choice for their application.

  In this chapter, random-phase multisines are used since this is a reasonable choice for systems that are dominantly linear~\TODO{ref: BLA}.
  Hence, $\phi_k$ is the outcome of a uniform random process over $\left[0,2\pi\right[$.
  This leaves two sets of free parameters to create a signal that suits our needs:
  \begin{itemize}
    \item the amplitude of the spectral lines $A_k$, and
    \item the frequency grid $\left\{f_1, \ldots, f_F \right\}$.
  \end{itemize}
  Note that the $k^{\text{th}}$ term in \eqref{eq:excitation:MultisineCT} is periodic with period $T_k \isdef f_k^{-1}$.

@@

  As in \citep{Pintelon2012, Gallager2008, matrixcookbook}, and equivalently to the derivation in \secref{sec:excitation:inputDesign}, the covariance matrix $\C_{G\left(\omega_{\interest}\right)}$ of the parametric transfer function $G(\theta,\omega)$ can be shown to be approximately equal to
  \begin{align}
    \C_{G\left( \omega_{\interest}\right) } \approx
      \InnerQuadraticForm[\HT]
                         {\jac{\interest}}
                         {\FisherInformation_{\theta}^{-1}}
     \label{eq:excitation:varPeak}
  \end{align}
  with 
  \begin{equation}
    \FisherInformation_{\theta} =
    2
    \real{
          \InnerQuadraticForm[\HT]
                              {\jac{\excited}}
                              {\C_{X}^{-1}}
          }
    \label{eq:excitation:FisherInfo}
  \end{equation}
  the Fisher information matrix \citep{Pintelon2012}.

@@

  Instead of studying the response of all possible second-order sub-systems for every possible logarithmically spaced multisine-- a six-dimensional problem with parameters $(\omega_0, \alpha)$ for the grid, and $(\wz, \wn, \damping, \dc)$ for the sub-system-- we argue to simplify the problem in such a way that the results are still valid for the more complicated setting.

  \subsubsection{Exploiting the grid misalignment between the system and the grid}
  Denote $G(\omega; \wn)$ a second-order system as in \eqref{eq:excitation:secondOrderSystemCT} with natural frequency $\wn$, damping ratio $\damping$ and \gls{DC} gain $\dc$.
  Similarly, $G(\omega; m\wn)$ where $m \in \RealNumbersWithoutZero$ is a second order system that shares the damping and \gls{DC} gain of the previous system but has a natural frequency $m\wn$.
